Percussive Sound Generation with Virtual Listeners and Modular Synthesizers

### 2020-08-17 templates_aimusic2020/paper.tex
Updates from Overleaf

In this context, \textit{sampling rate}, is an important feature of digital sound, referring to the number of samples per second of audio (measured in Hertz, or Hz). Sampling rate is not the only important factor when recording audio as it is important to record the samples with not just speed, but also precision. Assuming perfect sensors, precision is the range of possible values we can assign to each sample. It is determined by bit depth: the number of bits we have to represent the values of each sample.  Today, standard quality audio often refers to sampling rates of 44.1 kHz and 48 kHz and bit depth of 16 (that's $2^{16}$ values), while \enquote{high quality} audio indicates an increase in bit rate or bit depth~\cite{reiss2016meta}. Although subject to diminishing returns, high quality audio may be preferable to most musicians and audio-engineers; In a meta-analysis of digital sound perception, Reiss found a small but statistically significant portion of people are able to discriminate the effect of standard and high quality audio with no prior training, and a dramatically higher detection rate after extensive training~\cite{reiss2016meta}.

\subsection{Loudness, Amplitudes and Envelopes}
Loudness is a subjective description of a sound's intensity or energy levels. It varies based on the complexity of sounds, the frequencies present, and hearing ability of the listener~\cite{fletcher1933loudness,cook1999chap6}. It can only be measured relatively, by establishing a benchmark sound and surveying populations on the relative intensities~\cite{cook1999chap6}. Since loudness and intensity of sound correlate with the amplitude of digital waveforms (the values assigned to the samples), an imperfect but convenient alternative method for inferring the the loudness of digital sounds is to compare relative amplitudes.

The peak level of a digital sound corresponds to the sample with the highest value. This method has its benefits, but leaves out an important factor: the duration of intensity. A more commonly used method of level detection is the Root Mean Square (RMS) method, which averages the values of a short sequence of samples for its calculation of levels.

Sounds typically vary in intensity as they unfold. This change in intensity is often described by the \enquote{envelope} of the sound, particularly in shorter samples. For digital sounds, Mitchell describes the envelope as either (since samples typically take the range of -1 to 1) of the borders that are created by graphing a signal and connecting the local absolute peak values~\cite{mitchell2009basicsynthChap6}. The envelope is generally described via 4 features: Attack, Decay, Sustain and Release (ADSR). Attack describes how quickly the peak loudness is reached. Decay for how quickly the sound drops to sustain level. Sustain is the duration of sustaining intensity (e.g how long a finger is kept on a piano key). Release describes the speed of fading to silence (e.g once a piano key is released).

Envelopes can be mathematically described and used to shape signals. A common approach in digital sound synthesis is to output all samples at a consistent amplitude and apply an envelope later down the synthesis chain. Digital and analog synthesizers often have built-in ADSR modules to shape the volume of the output and other parameters.

[Graph of envelope shaping a signal]
% note: make sure you're using terms signal and waveform properly
\subsection{Frequency, Pitch and Spectrograms}
\textit{Frequency} is used to describe number of repetitions within a time-frame, or how frequently a cycle is repeated. The frequency of an audio signal is often measured in unit of hertz (cycles per second). Most sounds, particularly those from none virtual sources, are a combination of multiple different pressure waves with different frequencies and amplitudes. \textit{Pitch}, is a perceptual property of sound that is tied to the frequencies present in each sound. How we perceive and describe the pitch of a sound (e.g high-pitched vs low-pitched) is heavily dependent on the characteristics (frequency, amplitude, duration, etc) of the waveforms it contains. Some sounds such as piano chords or pure tones have easily discernable pitches. Others, such as \enquote{pink noise} or the sound of rain do not. Yet another factor to consider is the hearing ability of the subject, which varies between people based on factors such as age, environment, musical training, etc~\cite{reiss2016meta,alain2007age,newman2012grm7}.

Spectrograms are graphs used to showcase and analyze the duration and amplitude of frequencies present in a sound. Since we often do not have access to the time-variant systems behind sounds of interest, it is common to use function decomposition methods to break sounds down to a combination of simpler functions. A commonly used method for the breakdown of complex, time-variant functions is the fourier transformation and its variations. One such method is the discrete fourier transformation (DFT) and its inverse, often used to convert digital sound from its time domain representation (sequence of samples) to its frequency domain representation (sets of frequency ranges and their amplitude) and vice versa.\\\\

[spectrogram figures]
[figure of this breakdown]

>>>>

\subsection{Frequency, Pitch and Spectrograms}
\textit{Frequency} is used to describe number of repetitions within a time-frame, or how frequently a cycle is repeated. The frequency of an audio signal is often measured in unit of hertz (cycles per second). Most sounds, particularly those from none virtual sources, are a combination of multiple different pressure waves with different frequencies and amplitudes. \textit{Pitch}, is a perceptual property of sound that is tied to the frequencies present in each sound. How we perceive and describe the pitch of a sound (e.g high-pitched vs low-pitched) is heavily dependent on the characteristics (frequency, amplitude, duration, etc) of the waveforms it contains. Some sounds such as piano chords or pure tones have easily discernable pitches. Others, such as \enquote{pink noise} or the sound of rain do not. Yet another factor to consider is the hearing ability of the subject, which varies between people based on factors such as age, environment, musical training, etc~\cite{reiss2016meta,alain2007age,newman2012grm7}.
